Privacy-Preserving Localisation Using Private Linear-Combination Aggregation

### 2020-11-05 main.tex
Continued with template matching Cleaned text from template comments and many grammarly errors

% autosam.tex
% Annotated sample file for the preparation of LaTeX files
% for the final versions of papers submitted to or accepted for 
% publication in AUTOMATICA.

% See also the Information for Authors.

% Make sure that the zip file that you send contains all the 
% files, including the files for the figures and the bib file.

% Output produced with the elsart style file does not imitate the
% AUTOMATICA style. The style file is generic for all Elsevier
% journals and the output is laid out for easy copy editing. The
% final document is produced from the source file in the
% AUTOMATICA style at Elsevier.

% You may use the style file autart.cls to obtain a two-column 
% document (see below) that more or less imitates the printed 
% Automatica style. This may helpful to improve the formatting 
% of the equations, tables and figures, and also serves to check 
% whether the paper satisfies the length requirements.

% Please note: Authors must not create their own macros.

% For further information regarding the preparation of LaTeX files 
% for Elsevier, please refer to the "Full Instructions to Authors" 
% from Elsevier's anonymous ftp server on ftp.elsevier.nl in the
% directory pub/styles, or from the internet (CTAN sites) on
% ftp.shsu.edu, ftp.dante.de and ftp.tex.ac.uk in the directory
% tex-archive/macros/latex/contrib/supported/elsevier.

% The use of LaTeX2e is preferred.
%\documentclass{elsart}

% Enable this line and disable the 
% preceding line to obtain a two-column 
% document whose style resembles the
% printed Automatica style.
\documentclass[twocolumn]{autart}

% Include this line if your 
% document contains figures,
\usepackage{graphicx}
% or this line, depending on which
% you prefer.
%\usepackage[dvips]{epsfig}

% User imported packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{tikz}

% Local packages
\usepackage{ISASmacros/isasmathmacros}

@@

\begin{document}

\begin{frontmatter}
% Running title for regular 
% papers but only if the title  
% is over 5 words. Running title 
% is not shown in output.
%\runtitle{Insert a suggested running title}

% Title, preferably not more than 10 words.
\title{Privacy-Preserving Localisation Using Private Linear-Combination Aggregation\thanksref{footnoteinfo}}

\thanks[footnoteinfo]{This paper was not presented at any IFAC 
meeting. Corresponding author Uwe~D.~Hanebeck. Tel. +49--721--608--43909.}

% Add the 
% e-mail address 
% (ead) as shown
\author[ISAS]{Marko Ristic}\ead{marko.ristic@kit.edu},
\author[ISAS]{Benjamin Noack}\ead{noack@kit.edu},
\author[ISAS]{Uwe D. Hanebeck}\ead{uwe.hanebeck@kit.edu}

% Please supply
\address[ISAS]{Intelligent Sensor-Actuator-Systems Laboratory, Institute for Anthropomatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}

% Five to ten keywords,  
% chosen from the IFAC 
% keyword list or with the 
% help of the Automatica 
% keyword wizard
\begin{keyword}
System state estimation; Data privacy; Sensor fusion; Kalman filters.
\end{keyword}

@@

% Abstract of not more than 200 words.
\begin{abstract}
Distributed state estimation and localisation methods have become increasingly popular with the rise of ubiquitous computing, and have led naturally to an increased concern regarding data and estimation privacy. Traditional distributed sensor navigation methods involve the leakage of sensor or navigator information during localisation protocols, thus not preserving participants’ data privacy. Existing approaches which provide such guarantees fail to address sensor and navigator privacy in some common non-linear measurement model applications and forfeit broad applicability. We define the novel notion of linear-combination aggregation encryption and provide a cryptographically secure instance that is applied to the Extended Information Filter with range-sensor measurements. Our method keeps navigator location, sensor locations and sensor measurements private during navigation, and is implemented and run on simulations to evaluate its accuracy and performance. The novel computationally plausible and provably secure range-based localisation filter has direct applications to environments where nodes are not trusted or data is considered sensitive.
\end{abstract}

\end{frontmatter}

@@

\section{Introduction}
Localisation methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,liContributedReviewSourcelocalization2016}. Ongoing advancements in portable computing power and sensor capabilities have led to the development of various forms of localisation methods, from range-free and range-based signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes}, acoustic range estimates \cite{beutlerNewNonlinearFiltering2004} and magnetic field measurements \cite{sieblerLocalizationMagneticField2020} to the TDOA and TOA pseudo-range multilateration methods for aircraft \cite{liContributedReviewSourcelocalization2016}. Range-based localisation methods, including the use of radar measurements, GPS clocks and signal strength measurements, have found particularly large application due to their robustness and accuracy. These methods use distance measurements, or computed estimates, to produce location estimates and typically involve gathering sensor details such as measurement noise, geographical location and observations centrally. Broadcasting or gathering information in this way, however, implies a trust between participants within a trusted communication network to ensure shared data may not be used for individual malicious gain. With the advancements in cloud and ubiquitous computing, changes in possible use-cases involving distributed sensors have made the requirements of information security and user privacy more apparent than before \cite{brennerSecretProgramExecution2011,renSecurityChallengesPublic2012}. For example, an aircraft localisation scenario in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Or in autonomous vehicle state estimation, benefit from neighbouring vehicles' information may be desired while no participant wishes to share information about their hardware. Additionally, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations. All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data. Our contributions in this work will be presented in two parts. We first propose a novel aggregation-based encryption scheme which allows private aggregation of weighted sums, before we develop a localisation filter using this scheme such that sensor and navigator privacy is preserved.

@@

\begin{description}

@@

\end{description}

@@

\begin{rem}
    From the aggregation definition above and the distributive property of real numbers, it is clear that an alternative method to the bi-directional communication requirement would be computing $\sum^{m}_{j=1}\omega_j^{(t)}\sum^{n}_{i=1} x_{i,j}^{(t)}$ by the navigator, requiring only uni-directional communication by sending values $x_{i,j}^{(t)}, 1 \leq j \leq m$ from each sensor $i$. We justify the use of bi-directional communication by reducing communication costs when the number of weights is larger than the number of sensors, $m>n$, and by sending fewer weights in the presence of repeats, as will be shown to be the case in Section \ref{subsec:algorithm}.
\end{rem}

@@

\subsection{Participant Capabilities} \label{subsec:capabilities}
Before we can define concrete security requirements and a filtering protocol, we require some assumptions on the capabilities of the navigator and sensors in our problem scenario.
\begin{description}
    \item[Global navigator broadcast] From the decomposition defined above, we know that bi-directional communication between navigator and sensors is required. We make the assumption that broadcast information from the navigator is received by \textit{all} sensors involved in the protocol.

@@

    Misbehaving sensors are a known problem in estimation theory, often requiring inconsistency or outlier detection mechanisms \cite{lazosSeRLocSecureRangeindependent2004,ben-galOutlierDetection2005}. As privacy-preserving implementations of such mechanisms require additional complications, we will not consider the scenario in this work.
    \item[Computational capabilities] The computational requirements for computing encryptions and homomorphic operations are typically more costly than their plaintext equivalents. We make the implicit assumption that all involved parties are computationally capable of computing the required encryption and filter operations.
\end{description}
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[font=\footnotesize]
    % Sizes
    \tikzmath{\srange = 3.54; \sradius = 0.04;}
    % Sensor locations (sensor 3 is shared)
    \coordinate (s1) at (0.8523,-4.1508) {};
    \coordinate (s2) at (3.507,-5.5587) {};
    \coordinate (s3) at (2.5,-2.75) {} {};
    \coordinate (s4) at (4,-1.25) {};
    \coordinate (s5) at (5.4346,-3.1274) {};
    % Navigator location
    \coordinate (nav_loc) at (3.5,-3.25) {};
    % Bounding rectangle
    \coordinate (brect_start) at (0,0);
    \coordinate (brect_end) at (6.5, -6.5);

@@

\subsection{Security Notions}
The security of a cryptographic scheme is typically defined by a security \textit{game}, which captures both the desired privacy guarantees, as well as the capabilities of attackers \cite{katzIntroductionModernCryptography2008}. The typical security notion for a homomorphic encryption scheme is Indistinguishability under Chosen Plaintext Attack (IND-CPA) \cite{chaseSecurityHomomorphicEncryption2017}. 
\begin{defn}
An encryption scheme meets IND-CPA security if an attacker who can choose plaintext messages to be encrypted at will, gains no additional information about an unknown plaintext message when they learn only its encryption.

The formal security game for IND-CPA has been given in Appendix \ref{app:ind_cpa}.
\end{defn}
Private aggregation schemes aim for the security notion of Aggregator Obliviousness (AO) \cite{shiPrivacyPreservingAggregationTimeSeries2011}. 
\begin{defn}
An encryption scheme meets AO security if no colluding subset of participants \textit{excluding} the aggregator gains additional information about the remaining aggregation values given only their encryptions, while any colluding subset \textit{including} the aggregator learns only their sum.

The formal security game for AO has been given in Appendix \ref{app:ao}.
\end{defn}

@@

\section{Private Linear-Combination Aggregation} \label{sec:lcao_definition}
As stated in Section \ref{subsec:problem_decomposition}, we decompose the localisation method such that the multi-party aggregation protocol, given in Figure \ref{fig:agg_steps}, captures all communication between the navigator and sensors. To perform this protocol in a secure manner, we must first describe the security properties we aim to achieve during aggregation. Broadly speaking, we want a cryptographic scheme which allows instances of homomorphically computed linear combinations of encrypted weights, to be summed by a private aggregation scheme. That is, we do not want sensors to learn the navigator weights, while we do not want the navigator to learn individual weighted linear combinations. This can be summarised by the two informal security notions:
\begin{description}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any additional knowledge about the navigator weights $\omega_j,\,1\leq j \leq m$ from receiving only their encryptions from the current and previous instances, and the ability to encrypt plaintexts of their choice.
    \item[Private Linear-Combination Aggregation] No colluding subset \textit{excluding} the aggregator gains additional information about the remaining sensor values to be weighted $x^{(t)}_{i,j},\,1\leq j\leq m$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $y_i$ from the current and previous instances. Any colluding subset \textit{including} the aggregator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}y_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} x^{(t)}_{i,j}\omega^{(t)}_j$.
\end{description}

\begin{rem} \label{rem:lcao_leakage}
    The notion of a leakage function including parameters from the aggregator requires extra care to be taken when giving its definition. Since an attacker may compromise the aggregator, they have control over the choice of these parameters, and therefore over the leakage function. We note that in the leakage function above, $\sum^n_{i=n}\sum^m_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learned by the colluding subset, \textit{e.g.} $\sum^m_{j=1}x^{(t)}_{i,j}$ given weights $(1,0,\dots,0)$, but that individual sensor values $x^{(t)}_{i,j}$ remain private due to the requirement that all sensors receive the same weights.
\end{rem}

@@

\begin{description}
    \item[$\mathsf{Setup}(\kappa)$] On input of security paramater $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the aggregator's public and private keys $pk_0$ and $sk_0$, and the user private keys $sk_i,\,1\leq i \leq n$.
    \item[$\mathsf{Enc}(pk_0, a)$] The aggregator and users can encrypt any value $a$ with the aggregator public key $pk_0$, and obtain the encryption $\mathcal{E}_{pk_0}(a)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1^{(t)}),\dots,\mathcal{E}_{pk_0}(\omega_m^{(t)}), x^{(t)}_{i,1},\dots,x^{(t)}_{i,m})$] At instance $t$, user $i$ computes and obtains the encrypted linear combination $y^{(t)}_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, y^{(t)}_1,\dots,y^{(t)}_n)$] At instance $t$, the aggregator computes the aggregation of linear combinations $\sum^{n}_{i=1}y_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} x^{(t)}_{i,j}\omega^{(t)}_j$ using its public and private keys $pk_0$, $sk_0$.
\end{description}

@@

\begin{description}
    \item[$\mathsf{Setup}(\kappa)$] On input parameter $\kappa$, generate two equal length, sufficiently large, primes $p$ and $q$, and compute $N=pq$. Define a hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}^*$, choose an $m>1$ as the number of weights to combine, and set public parameter $\mathsf{pub}=H$, aggregator public key $pk_0 = N$ and aggregator private key $sk_0=(p,q)$. The remaining user secret keys are generated by choosing $sk_i,\,1\leq i\leq n-1$ uniformly from $\mathbb{Z}_{N^2}$ and setting the last key as $sk_n = -\sum^{n-1}_{i=1}sk_i \pmod{N^2}$.

@@

    \item[$\mathsf{AggDec}(t, pk_0, sk_0, y^{(t)}_1,\dots,y^{(t)}_n)$] Aggregation is computed as $y^{(t)} = \prod^n_{i=1}y^{(t)}_i \pmod{N^2}$, removing aggregation noise terms, and is followed by Paillier decryption
    \begin{equation}
        \begin{split}
            \sum^{n}_{i=1}\sum^{m}_{j=1}&x^{(t)}_{i,j}\omega^{(t)}_j =\\
            &\frac{L((y^{(t)})^\lambda\pmod{N^2})}{L((N+1)^\lambda\pmod{N^2})} \pmod{N}\,.
        \end{split} \label{eqn:our_scheme_decrypt}
    \end{equation}
    The correctness of aggregation can be seen from
    \begin{align*}
        y^{(t)} &= \prod^n_{i=1}H(t)^{sk_i}(N+1)^{\sum^{m}_{j=1}x_{i,j}\omega_j}r_i^N \pmod{N^2} \\
        \begin{split}
            &= H(t)^{\sum^n_{i=1}sk_i}\cdot \\
            &\qquad\qquad \prod^n_{i=1}(N+1)^{\sum^{m}_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j}r_i^N \pmod{N^2}
        \end{split}\\
        &= (N+1)^{\sum^n_{i=1}\sum^{m}_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j}r'^N \pmod{N^2}\,,
    \end{align*}
    for some values $r_i,r' \in \mathbb{Z}_N,\,1\leq i \leq n$.
\end{description}

@@

\begin{rem} \label{rem:lcao_implicit_weight}
    The construction of this scheme additionally supports the linear combination of the implicit weight, $\omega^{(t)}_j = 1$, by replacing
    \begin{equation}
        \mathcal{E}_{pk_0}(1)^{x^{(t)}_{i,j}} = (N+1)^{1}r_j^N \pmod{N^2}
    \end{equation}
    in \eqref{eqn:our_scheme_lin_comb} by
    \begin{equation}
        (N+1)^{x^{(t)}_{i,j}} \pmod{N^2}\,,
    \end{equation}
    due to the removal of $r_j^N$ terms during decryption. This can be used to reduce the broadcast communication cost by one weight.
\end{rem}

@@

    \item[Update] The navigator aggregates measurement vectors and matrices before decrypting them. The typical EIF update equations can then be computed. This is shown in Algorithm \ref{alg:nav_update}.
\end{description}

>>>>

%% bare_jrnl_compsoc.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% Computer Society journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

\documentclass[10pt,journal,compsoc]{IEEEtran}

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
\else
  \usepackage[dvips]{graphicx}
\fi

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% 
% 8888888 888b     d888 8888888b.   .d88888b.  8888888b. 88888888888 .d8888b.  
%   888   8888b   d8888 888   Y88b d88P" "Y88b 888   Y88b    888    d88P  Y88b 
%   888   88888b.d88888 888    888 888     888 888    888    888    Y88b.      
%   888   888Y88888P888 888   d88P 888     888 888   d88P    888     "Y888b.   
%   888   888 Y888P 888 8888888P"  888     888 8888888P"     888        "Y88b. 
%   888   888  Y8P  888 888        888     888 888 T88b      888          "888 
%   888   888   "   888 888        Y88b. .d88P 888  T88b     888    Y88b  d88P 
% 8888888 888       888 888         "Y88888P"  888   T88b    888     "Y8888P"  
%                                                                              
%                                                                              
%                                                                              
%

% Graphics
\usepackage{tikz}
% Maths
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{mathtools}
% Algorithms
\usepackage{algorithm}
\usepackage{algpseudocode}
% Local
\usepackage{ISASmacros/isasmathmacros}
% Equal last columns
\usepackage{flushend}

% Environments
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{definition}
\newtheorem{theorem}{Theorem}[section]

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

@@

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Privacy-Preserving Localisation Using Private Linear-Combination Aggregation}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs

\author{Marko~Ristic,
        Benjamin~Noack,~\IEEEmembership{Member,~IEEE,}
        and~Uwe~D.~Hanebeck,~\IEEEmembership{Fellow,~IEEE}% <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem M. Ristic, B. Noack and U. D. Hanebeck are with the Intelligent Sensor-Actuator-Systems Laboratory, Institute for Anthropomatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany.\protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
E-mail: marko.ristic@kit.edu}% <-this % stops an unwanted space
}

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Distributed state estimation and localisation methods have become increasingly popular with the rise of ubiquitous computing, and have led naturally to an increased concern regarding data and estimation privacy. Traditional distributed sensor navigation methods involve the leakage of sensor or navigator information during localisation protocols, thus not preserving participants’ data privacy. Existing approaches which provide such guarantees fail to address sensor and navigator privacy in some common non-linear measurement model applications and forfeit broad applicability. We define the novel notion of linear-combination aggregation encryption and provide a cryptographically secure instance that is applied to the Extended Information Filter with range-sensor measurements. Our method keeps navigator location, sensor locations and sensor measurements private during navigation, and is implemented and run on simulations to evaluate its accuracy and performance. The novel computationally plausible and provably secure range-based localisation filter has direct applications to environments where nodes are not trusted or data is considered sensitive.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
State Estimation, Data Privacy, Sensor Fusion, Kalman Filters.
\end{IEEEkeywords}}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEtitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynontitleabstractindextext when the compsoc 
% or transmag modes are not selected <OR> if conference mode is selected 
% - because all conference papers position the abstract like regular
% papers do.
\IEEEdisplaynontitleabstractindextext
% \IEEEdisplaynontitleabstractindextext has no effect when using
% compsoc or transmag under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

@@

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
% Computer Society journal (but not conference!) papers do something unusual
% with the very first section heading (almost always called "Introduction").
% They place it ABOVE the main text! IEEEtran.cls does not automatically do
% this for you, but you can achieve this effect with the provided
% \IEEEraisesectionheading{} command. Note the need to keep any \label that
% is to refer to the section immediately after \section in the above as
% \IEEEraisesectionheading puts \section within a raised box.

\IEEEPARstart{L}{ocalisation} methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,liContributedReviewSourcelocalization2016}. Ongoing advancements in portable computing power and sensor capabilities have led to the development of various forms of localisation methods, from range-free and range-based signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes}, acoustic range estimates \cite{beutlerNewNonlinearFiltering2004} and magnetic field measurements \cite{sieblerLocalizationMagneticField2020} to the TDOA and TOA pseudo-range multilateration methods for aircraft \cite{liContributedReviewSourcelocalization2016}. Range-based localisation methods, including the use of radar measurements, GPS clocks and signal strength measurements, have found particularly large application due to their robustness and accuracy. These methods use distance measurements, or computed estimates, to produce location estimates and typically involve gathering sensor details such as measurement noise, geographical location and observations centrally. Broadcasting or gathering information in this way, however, implies a trust between participants within a trusted communication network to ensure shared data may not be used for individual malicious gain. With the advancements in cloud and ubiquitous computing, changes in possible use-cases involving distributed sensors have made the requirements of information security and user privacy more apparent than before \cite{brennerSecretProgramExecution2011,renSecurityChallengesPublic2012}. For example, an aircraft localisation scenario in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Or in autonomous vehicle state estimation, benefit from neighbouring vehicles' information may be desired while no participant wishes to share information about their hardware. Additionally, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations. All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data. Our contributions in this work will be presented in two parts. We first propose a novel aggregation-based encryption scheme which allows private aggregation of weighted sums, before we develop a localisation filter using this scheme such that sensor and navigator privacy is preserved.

@@

\begin{LaTeXdescription}

@@

\end{LaTeXdescription}

@@

\begin{remark}
    From the aggregation definition above and the distributive property of real numbers, it is clear that an alternative method to the bi-directional communication requirement would be computing $\sum^{m}_{j=1}\omega_j^{(t)}\sum^{n}_{i=1} x_{i,j}^{(t)}$ by the navigator, requiring only uni-directional communication by sending values $x_{i,j}^{(t)}, 1 \leq j \leq m$ from each sensor $i$. We justify the use of bi-directional communication by reducing communication costs when the number of weights is larger than the number of sensors, $m>n$, and by sending fewer weights in the presence of repeats, as will be shown to be the case in Section \ref{subsec:algorithm}.
\end{remark}

@@

\subsection{Participant Capabilities} \label{subsec:capabilities}
Before we can define concrete security requirements and a filtering protocol, we require some assumptions on the capabilities of the navigator and sensors in our problem scenario.
\begin{LaTeXdescription}
    \item[Global navigator broadcast] From the decomposition defined above, we know that bi-directional communication between navigator and sensors is required. We make the assumption that broadcast information from the navigator is received by \textit{all} sensors involved in the protocol.

@@

    Misbehaving sensors are a known problem in estimation theory, often requiring inconsistency or outlier detection mechanisms \cite{lazosSeRLocSecureRangeindependent2004,ben-galOutlierDetection2005}. As privacy-preserving implementations of such mechanisms require additional complications, we will not consider the scenario in this work.
    \item[Computational capabilities] The computational requirements for computing encryptions and homomorphic operations are typically more costly than their plaintext equivalents. We make the implicit assumption that all involved parties are computationally capable of computing the required encryption and filter operations.
\end{LaTeXdescription}
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[font=\footnotesize]
    % Sizes
    \tikzmath{\srange = 3.54; \sradius = 0.04;}
    % Sensor locations (sensor 3 is shared)
    \coordinate (s1) at (0.8523,-4.1508) {};
    \coordinate (s2) at (3.507,-5.5587) {};
    \coordinate (s3) at (2.5,-2.75) {} {};
    \coordinate (s4) at (4,-1.25) {};
    \coordinate (s5) at (5.4346,-3.1274) {};
    % Navigator location
    \coordinate (nav_loc) at (3.5,-3.25) {};
    % Bounding rectangle
    \coordinate (brect_start) at (0,0);
    \coordinate (brect_end) at (6.5, -6.5);

@@

\subsection{Security Notions}
The security of a cryptographic scheme is typically defined by a security \textit{game}, which captures both the desired privacy guarantees, as well as the capabilities of attackers \cite{katzIntroductionModernCryptography2008}. The typical security notion for a homomorphic encryption scheme is Indistinguishability under Chosen Plaintext Attack (IND-CPA) \cite{chaseSecurityHomomorphicEncryption2017}. 
\begin{definition}
An encryption scheme meets IND-CPA security if an attacker who can choose plaintext messages to be encrypted at will, gains no additional information about an unknown plaintext message when they learn only its encryption.

The formal security game for IND-CPA has been given in Appendix \ref{app:ind_cpa}.
\end{definition}
Private aggregation schemes aim for the security notion of Aggregator Obliviousness (AO) \cite{shiPrivacyPreservingAggregationTimeSeries2011}. 
\begin{definition}
An encryption scheme meets AO security if no colluding subset of participants \textit{excluding} the aggregator gains additional information about the remaining aggregation values given only their encryptions, while any colluding subset \textit{including} the aggregator learns only their sum.

The formal security game for AO has been given in Appendix \ref{app:ao}.
\end{definition}

@@

\section{Private Linear-Combination Aggregation} \label{sec:lcao_definition}
As stated in Section \ref{subsec:problem_decomposition}, we decompose the localisation method such that the multi-party aggregation protocol, given in Figure \ref{fig:agg_steps}, captures all communication between the navigator and sensors. To perform this protocol in a secure manner, we must first describe the security properties we aim to achieve during aggregation. Broadly speaking, we want a cryptographic scheme which allows instances of homomorphically computed linear combinations of encrypted weights, to be summed by a private aggregation scheme. That is, we do not want sensors to learn the navigator weights, while we do not want the navigator to learn individual weighted linear combinations. This can be summarised by the two informal security notions:
\begin{LaTeXdescription}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any additional knowledge about the navigator weights $\omega_j,\,1\leq j \leq m$ from receiving only their encryptions from the current and previous instances, and the ability to encrypt plaintexts of their choice.
    \item[Private Linear-Combination Aggregation] No colluding subset \textit{excluding} the aggregator gains additional information about the remaining sensor values to be weighted $x^{(t)}_{i,j},\,1\leq j\leq m$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $y_i$ from the current and previous instances. Any colluding subset \textit{including} the aggregator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}y_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} x^{(t)}_{i,j}\omega^{(t)}_j$.
\end{LaTeXdescription}

\begin{remark} \label{rem:lcao_leakage}
    The notion of a leakage function including parameters from the aggregator requires extra care to be taken when giving its definition. Since an attacker may compromise the aggregator, they have control over the choice of these parameters, and therefore over the leakage function. We note that in the leakage function above, $\sum^n_{i=n}\sum^m_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learned by the colluding subset, \textit{e.g.} $\sum^m_{j=1}x^{(t)}_{i,j}$ given weights $(1,0,\dots,0)$, but that individual sensor values $x^{(t)}_{i,j}$ remain private due to the requirement that all sensors receive the same weights.
\end{remark}

@@

\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input of security paramater $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the aggregator's public and private keys $pk_0$ and $sk_0$, and the user private keys $sk_i,\,1\leq i \leq n$.
    \item[$\mathsf{Enc}(pk_0, a)$] The aggregator and users can encrypt any value $a$ with the aggregator public key $pk_0$, and obtain the encryption $\mathcal{E}_{pk_0}(a)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1^{(t)}),\dots,\mathcal{E}_{pk_0}(\omega_m^{(t)}), x^{(t)}_{i,1},\dots,x^{(t)}_{i,m})$] At instance $t$, user $i$ computes and obtains the encrypted linear combination $y^{(t)}_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, y^{(t)}_1,\dots,y^{(t)}_n)$] At instance $t$, the aggregator computes the aggregation of linear combinations $\sum^{n}_{i=1}y_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} x^{(t)}_{i,j}\omega^{(t)}_j$ using its public and private keys $pk_0$, $sk_0$.
\end{LaTeXdescription}

@@

\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input parameter $\kappa$, generate two equal length, sufficiently large, primes $p$ and $q$, and compute $N=pq$. Define a hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}^*$, choose an $m>1$ as the number of weights to combine, and set public parameter $\mathsf{pub}=H$, aggregator public key $pk_0 = N$ and aggregator private key $sk_0=(p,q)$. The remaining user secret keys are generated by choosing $sk_i,\,1\leq i\leq n-1$ uniformly from $\mathbb{Z}_{N^2}$ and setting the last key as $sk_n = -\sum^{n-1}_{i=1}sk_i \pmod{N^2}$.

@@

    \item[$\mathsf{AggDec}(t, pk_0, sk_0, y^{(t)}_1,\dots,y^{(t)}_n)$] Aggregation is computed as $y^{(t)} = \prod^n_{i=1}y^{(t)}_i \pmod{N^2}$, removing aggregation noise terms, and is followed by Paillier decryption
    \begin{equation}
        \begin{split}
            \sum^{n}_{i=1}\sum^{m}_{j=1}&x^{(t)}_{i,j}\omega^{(t)}_j =\\
            &\frac{L((y^{(t)})^\lambda\pmod{N^2})}{L((N+1)^\lambda\pmod{N^2})} \pmod{N}\,.
        \end{split} \label{eqn:our_scheme_decrypt}
    \end{equation}
    The correctness of aggregation can be seen from
    \begin{align*}
        y^{(t)} &= \prod^n_{i=1}H(t)^{sk_i}(N+1)^{\sum^{m}_{j=1}x_{i,j}\omega_j}r_i^N \pmod{N^2} \\
        \begin{split}
            &= H(t)^{\sum^n_{i=1}sk_i}\cdot \\
            &\qquad\qquad \prod^n_{i=1}(N+1)^{\sum^{m}_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j}r_i^N \pmod{N^2}
        \end{split}\\
        &= (N+1)^{\sum^n_{i=1}\sum^{m}_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j}r'^N \pmod{N^2}\,,
    \end{align*}
    for some values $r_i,r' \in \mathbb{Z}_N,\,1\leq i \leq n$.
\end{LaTeXdescription}

@@

\begin{remark} \label{rem:lcao_implicit_weight}
    The construction of this scheme additionally supports the linear combination of the implicit weight, $\omega^{(t)}_j = 1$, by replacing
    \begin{equation}
        \mathcal{E}_{pk_0}(1)^{x^{(t)}_{i,j}} = (N+1)^{1}r_j^N \pmod{N^2}
    \end{equation}
    in \eqref{eqn:our_scheme_lin_comb} by
    \begin{equation}
        (N+1)^{x^{(t)}_{i,j}} \pmod{N^2}\,,
    \end{equation}
    due to the removal of $r_j^N$ terms during decryption. This can be used to reduce the broadcast communication cost by one weight.
\end{remark}

@@

    \item[Update] The navigator aggregates measurement vectors and matrices before decrypting them. The typical EIF update equations can then be computed. This is shown in Algorithm \ref{alg:nav_update}.
\end{LaTeXdescription}
